---
title: "SNAP Data 1990-2024"
format: 
    html:
      code-fold: true 
  
theme: mycustom.scss

editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(highcharter)
library(sjlabelled)

source("functionsg.R")

snap_final <- read_rds("data/snap_df3.rds")

# library(showtext)
# font_add_google(name = "Montserrat", family = "Montserrat")
# showtext_auto()


extrafont::loadfonts(quiet = FALSE)

mysource <- "Source: U.S. Department of Agriculture 2024"

fig_basic <- snap_final %>% 
  filter(state_abbrev == "US") %>% 
  rename("Number of households" = snaphh2, 
         date = ym) %>% 
  ggplot(aes(x=dat, y= `Number of households`))+
  geom_col(alpha = 0.5, fill = nord_silver[5])+
  custom_style3()+
  scale_x_date(name = NULL)+
  scale_y_continuous(name = NULL, 
                   labels = scales::comma)+
  labs(title = "Number of US Households Receiving SNAP Benefits", 
       caption = mysource)

  
# plotly::ggplotly(fig_basic) %>% 
#   plotly::layout(font = list(family = "Montserrat"), 
#                  title = list(x = 0.4,            # x position with respect to `xref` in normalized 
#                                         # coordinates from "0" (left) to "1" (right).
#                       xanchor ="right"), 
#                  xaxis = list(title = ''), 
#                  yaxis = list(title = ''), 
#                   margin = list(
#       l = 0,
#       r = 0,
#       b = 0,
#       t = 40, pad = 1
#     ))

```

## Number of households receiving benefits

In September 2024, more than 22.5 million households receive SNAP benefits. The number of households receiving assistance has stayed above 21 million since April 2020. The current levels are just below the record amount between 2012 and 2014.

```{r}

snap_final %>% 
  filter(state_abbrev == "US") %>% 
  hchart(., "column",
         hcaes(x = ym, y = round(snaphh2)),
         color = nord_silver[5], 
         name = "Number of households") %>% 
  hc_title(text = "Number of US Households Receiving SNAP Benefits") %>% 
  hc_xAxis(title = list(text = "")) %>% 
  hc_yAxis(title = list(text = "")) %>% 
  hc_caption(text = "Source: U.S. Department of Agriculture") %>% 
  hc_add_theme(hc_theme_cstm())

```

## Benefit amounts over time

Current benefits average around \$350 per household and about \$190 per person. Benefit amounts peaked during the more generous Covid relief period in Nov 2022 averaging \$260 per person and \$490 per household.

```{r}


snap_final %>% 
  filter(state_abbrev == "US") %>% 
  hchart(., "line",
         hcaes(x = ym, y= benperhh), 
         color = nord_silver[5], 
         name = "Benefit amount per household", 
         tooltip = list(valueDecimals = 0, valuePrefix = "$")) %>% 
  hc_add_series(data = snap_final %>%
                  filter(state_abbrev == "US"),
                "line", 
         hcaes(x = ym, y = round(benperp)),
         color = nord_silver[1],
         name = "Benefit amount per person", 
         tooltip = list(valueDecimals = 0, valuePrefix = "$")) %>% 
  hc_title(text = "Benefit Amounts per Household and per Person") %>%
  hc_xAxis(title = list(text = "")) %>% 
  hc_yAxis(title = list(text = "")) %>% 
  hc_caption(text = "Source: U.S. Department of Agriculture") %>% 
  hc_add_theme(hc_theme_cstm())









```

## Comparisons at the state and regional levels

::: column-body-outset
```{r, fig.width=10, fig.height=8}


fig2 <- snap_final %>% 
  filter(!is.na(region_name)) %>% 
  ggplot(aes(x = ym, y = pct_snapp3))+
  geom_line(aes(group = state_abbrev),
            alpha = 0.3, 
             col = Maizgray) +
  geom_smooth(method = "loess", col = nord_silver[5])+
  facet_wrap(~region_name, 
             nrow = 1)+
  custom_style3()+
  scale_y_continuous(name = NULL, 
                     labels = add_pct_sign)+
  scale_x_date(name = NULL)+
  theme(plot.background = element_rect(fill = "#F0F0F0"), 
        strip.background = element_rect(color =  "#bfbfbf",
                                        fill = "#bfbfbf"), 
        strip.text = element_text(family = "Montserrat",
                                  hjust = 0.5, size = 14), 
        plot.subtitle = element_text(size = 15))+
  labs(title = "Percent of population receiveing SNAP Benefits by US State and Region", 
       subtitle = str_wrap("Each gray line shows a state in a US region while the blue line shows the average for each region.\n 
                           Relative to other regions, people in the south are more likely to receive SNAP benefits. Despite near record high numbers of people that receive SNAP benefits, the percentage of people that receive benefits has been on the decline since about 2014.", 
                           width = 95), 
       caption = "Source: U.S. Department of Agriculture")


fig2

```
:::
